Mg Depleted and K Enriched Stars from LAMOST Spectral Data

### 2018-01-10 article/ms.tex
Included python plotting script, latest images, changes to intro and meth
Note: the python plotting script won't work out of the box, needs to be in the appropriate location.

\section{Introduction}
The Mg-K anti correlation initially established in NGC 2419 continues to elude satisfactory explanation. Beyond the intrinsic scientific value of identifying the polluter object which is (presumably) responsible for puzzling Mg-K anti-correlation, a satisfactory explanation for this signature could potentially offer insight into the consistent under-abundance of K predicted by models \citep{kobayashi2011}. The signature also has well established links to globular-cluster evolution models.

NGC 2419 is the Milky Way's third most massive cluster, and one of the oldest at an age of 12.3 $\pm$ 1.3 Gy \citep{forbesandbridges2010}. It resides in the outer halo 87.5 kpc (original source unknown) from the galactic centre, and has been identified as perhaps one of the most unusual clusters in the galaxy.

A spectroscopic study of RGB giant stars in NGC 2419 led to the discovery discovery of a strong anti-correlation between Mg and K in around 30-40 \% of the studied stars, along with weaker relations in Si, Sc, Ca, Ti and V (\citet{mucciarelli2012}, \citet{cohenkirby2012}). Depletions of Mg within globular clusters are not uncommon, but they are always accompanied with an increase in [Fe/H], with the mechanism for this phenomena being the introduction of $\alpha$ element poor but iron rich material released from Type Ia supernovae locally polluting the 'normal' [Mg/Fe] material released from Type II supernovae. However, the two sub-populations are indistinguishable in their [Fe/H] abundance \citep{cohenkirby2012}, and this mechanism fails to account for the enhanced K abundance or trends in the other elements.

@@

With the exception of NGC 2808, in which all 4 of its known Mg depleted stars have been found to have an anti-correlation with K \citep{mucciarelli2015}, this signature has thus far been found to be contained within NGC 2419. A  targeted search looking [K/Fe] in RGB stars from clusters NGC 6752, NGC 6121, NGC 1904, and ω Centauri as well as 21 field stars found in all cases K abundances falling within the range of the Mg normal population in NGC 2419 \citep{carretta2013}.

@@

While it is plausible that the increased mass and greater distance from the gaseous disk of the Milky Way would cause a clusters first generation polluter stars to have an increased effect on the second generation of stars, this effect applies to all ejecta (assuming comparable energies). So while it might be expected that second generation stars clusters similar to NGC 2419 are present and are comprised of more material processed by the first generation than other lighter globular clusters, it doesn't explain why the ejecta from the first generation was apparently dominated by this as yet unknown polluter object resulting in the Mg-K anti-correlation. In the opinion of the author, it is the identity and physical characterisation of this unknown polluter that is of the greatest scientific interest.

An early modelling attempt at replicating the anomalous abundances in NGC 2419 assumed hot-bottom burning in AGB and super-AGB (SAGB) stars was able to reproduce the Mg-K anti-correlation, using the nuclear reaction pathway \ce{^{36}Ar(p,\gamma)^{37}K(\beta ^+ \nu)^{37}Ar(e^-,\nu)^{37}Cl(p,\gamma)^{38}Ar(p,\gamma)^39K} (corrected as per \cite{iliadis2016}) for nuclear synthesis. However, they were only able to do so if the reaction cross section was set to 100 times the commonly accepted rate or temperatures could be made higher than currently accepted values \citep{ventura2012}, achieving a burning temperature of 150 MK.

A more recent, comprehensive attempt to replicate not only the Mg and K abundances in NGC 2419 but also the abundances of Si, Sc, Ca, Ti and V, with the aim of constraining the temperatures and pressures required was undertaken by \cite{iliadis2016}. It is perhaps noteworthy that they identified a diferent main reaction pathway to that of \cite{ventura2012}, which is relegated to a minor minor seccondary pathway. The main pathway they identify is \ce{^{36}Ar(p,\gamma)^{37}K(\beta ^+ \nu)^{37}Ar(p,\gamma)^{38}K(\beta ^+ \nu)^{38}Ar(p,\gamma)^39K}. It was found that (with the exception of extremely high densities > $10^8$ g/cm$^3$) temperatures between 100 and 200 MK and densities between 10$^{-4}$ and 10$^8$ g/cm$^3$. Using this information, core and shell burning of low mass, high mass and super massive stars were ruled out as potential poluter candidates, as well as regular AGB stars. Super-AGB (SAGB) stars however were regarded as potential candidates, with only a relatively small (roughly 10-20 MK) increase in temperature required to fall in the acceptable band of parameter space identified. The other potential candidate identified was Novae, although the lack of detailed models of white dwarfs accreting material similar to NGC 2419 either from a binary companion or the intra cluster medium limit appraisals of its feasibility regarding whether enough ejecta could be produced and retained by the cluster. Based on current novae frequency in globular clusters determined by \cite{kato2013novae}, \cite{iliadis2016} conclude that the amount of material that would be produced by Novae is at most 1 \% of the total required mass to pollute 30 \% of NGC 2419.

@@

The data set of 454180 LAMOST giants was prepared by \citet{ho2017} for use with the supervised machine learning code \textit{The Cannon}. The LAMOST data was shifted to rest frame, normalised, and sampled to a common uniform wavelength grid between 3905 \AA and 9000 \AA. \textit{The Cannon} was used to infer $\text{T}_{\text{eff}}$, log(g), [Fe/H] and [$\alpha$/M] using 9952 common stars between LAMOST and the higher resolution APOGEE survey, through the use of a data-driven spectral model, referred to henceforth and simply 'the model'. For details regarding the process of the preparatory work, model generation and label transfer between the APOGEE and LAMOST surveys, the reader is directed to \citet{ho2017}. The data-driven model generated by \textit{The Cannon} is treated in this study as the typical spectra for each star.

@@

Three filters were applied independently, with the candidate pool formed from all unique stars identified each of the three filters. The first filter required the Mg 5184 line's Gaussian to satisfy $Amp > 0.05$ and $\frac{Amp}{Amp_{err}}>3$, and the K 7665 line to satisfy $Amp < 0.05$ and $\frac{Amp}{Amp_{err}}>3$. The second, more demanding filter was applied to require 2 of the three Mg lines to satisfy $Amp > 0.05$ and $\frac{Amp}{Amp_{err}}>3$, as well as either of the two K lines to satisfy $Amp < 0.05$ and $\frac{Amp}{Amp_{err}}>3$. The third filter removed the amplitude requirement to requiring only positivity on 2/3 Mg lines and negativity on \textit{both} K lines, in addition to data quality requirements $\chi^2<3$ and $S/N>30$
These filters together produced 384 unique candidate stars to be vetted manually.

The vetting process was done through visual examination of each case. The minimum criterion for a pass was two out of the three Mg lines to look at least plausibly depleted, and at least one of the K lines to look convincing. The apparent noisiness of the signal was qualitatively taken into account when making these determinations. Of the 384 candidates manually vetted, 188 were ultimately determined to be sufficiently strong candidates to be worth reporting in Table [INSERT TABLE REFERENCE HERE].

@@

\begin{figure}
\centering
\begin{subfigure}{.5\linewidth}
  \centering
  \includegraphics[width=.95\linewidth]{29277_Mg}
  \caption{Mg}
  \label{Mgposter}
\end{subfigure}%       %Note: the % to the left is VERY IMPORTANT.

\begin{subfigure}{.5\linewidth}
  \centering
  \includegraphics[width=.95\linewidth]{29277_K}
  \caption{K}
  \label{Kposter}
\end{subfigure}
\caption{}
\label{posterchildren}
\end{figure}

\begin{figure}
	\includegraphics[width=\columnwidth]{m_hhist.png}
    \caption{Metalicity Histogram}
    \label{mhist}
\end{figure}

\begin{figure}
	\includegraphics[width=\columnwidth]{teff_logg.png}
    \caption{Effective temperature, log surface gravity, metalicity}
    \label{mhist}
\end{figure}

\begin{figure}
	\includegraphics[width=\columnwidth]{galcoord.png}
    \caption{Galactic coordinates for Mg-K stars and known globular clusters}
    \label{mhist}
\end{figure}

>>>>

\section{Introduction}
The Mg-K anti correlation initially established in NGC 2419 continues to elude satisfactory explanation. Beyond the intrinsic scientific value of identifying the polluter object which is (presumably) responsible for the puzzling Mg-K anti-correlation, a satisfactory explanation for this signature could potentially offer insight into the underestimation of K abundance predicted by models \citep{kobayashi2011}. Such an explanation could also be beneficial to efforts to understand globular cluster formation and evolution.

NGC 2419 is the Milky Way's third most massive cluster, and one of the oldest at an age of 12.3 $\pm$ 1.3 Gy \citep{forbesandbridges2010}. It resides in the outer halo 87.5 kpc (original source unknown) from the galactic centre, and its chemical make-up makes it one of the most unusual clusters in the galaxy.

A spectroscopic study of RGB giant stars in NGC 2419 led to the discovery discovery of a strong anti-correlation between Mg and K in around 30-40 \% of the studied stars, along with weaker relations in Si, Sc, Ca, Ti and V (\citet{mucciarelli2012}, \citet{cohenkirby2012}). Depletions of Mg within globular clusters are not uncommon, but they are always accompanied with an increase in [Fe/H]. The mechanism for this phenomena is the introduction of $\alpha$ poor but iron rich material released from Type Ia supernovae locally polluting the 'normal' [Mg/Fe] material released from Type II supernovae. However, the two sub-populations in NGC 2419 are indistinguishable in their [Fe/H] abundance \citep{cohenkirby2012}. Further, Type Ia supernovae enrichment does not account for the enhanced K abundance or trends in the other elements.

@@

With the exception of NGC 2808, in which all 4 of its known Mg depleted stars have been found to have an anti-correlation with K \citep{mucciarelli2015}, this signature has thus far been found to be contained within NGC 2419. A  targeted search looking [K/Fe] in RGB stars from clusters NGC 6752, NGC 6121, NGC 1904, and $\omega$ Centauri as well as 21 field stars found in all cases K abundances falling within the range of the Mg normal population in NGC 2419 \citep{carretta2013}.

@@

While it is plausible that the increased mass and greater distance from the gaseous disk of the Milky Way would cause a clusters first generation polluter stars to have an increased effect on the second generation of stars, this effect would apply to all ejecta (assuming comparable energies). So while it might be expected that second generation stars in clusters similar to NGC 2419 made up of more material processed by the first generation than other lighter globular clusters, it doesn't explain why the ejecta from the first generation was apparently dominated by this as yet unknown polluter object resulting in the Mg-K anti-correlation. In the opinion of the author, it is the identity and physical characterisation of this unknown polluter that is of the greatest scientific interest, with such a result also hopefully answering the question as to the signatures uniqueness.

An early modelling attempt at replicating the anomalous abundances in NGC 2419 assumed hot-bottom burning in AGB and super-AGB (SAGB) stars \citep{ventura2012}. This study succeeded in reproducing the Mg-K anti-correlation, using nuclear reaction pathway \ce{^{36}Ar(p,\gamma)^{37}K(\beta ^+ \nu)^{37}Ar(e^-,\nu)^{37}Cl(p,\gamma)^{38}Ar(p,\gamma)^39K} (corrected as per \cite{iliadis2016}) for nuclear synthesis of K. However, the required abundance patterns consistent with NGC 2419 were only attainable if the reaction cross section was set to 100 times the commonly accepted rate or temperatures could be made higher than currently accepted values, achieving a burning temperature of 150 MK.

A more recent, comprehensive attempt to replicate the Mg and K abundances in NGC 2419 by \cite{iliadis2016} also attempted to reproduce the abundances of Si, Sc, Ca, Ti and V (elements reported as having weak correlations with Mg by \cite{cohenkirby2012}). The ultimate goal was to constrain the temperatures and densities required to produce the chemical signiature, and thereby provide insight into potential polluter candidates. It is perhaps noteworthy that they identified a diferent main reaction pathway to that of \cite{ventura2012}, which is relegated to a minor minor seccondary pathway. The main pathway they identify is \ce{^{36}Ar(p,\gamma)^{37}K(\beta ^+ \nu)^{37}Ar(p,\gamma)^{38}K(\beta ^+ \nu)^{38}Ar(p,\gamma)^39K}.

It was found that (with the exception of extremely high densities > $10^8$ g/cm$^3$) temperatures between 100 and 200 MK and densities between 10$^{-4}$ and 10$^8$ g/cm$^3$ were required. Using this information, core and shell burning of low mass, high mass and super massive stars were ruled out as potential poluter candidates, as well as regular AGB stars. Super-AGB (SAGB) stars however were regarded as potential candidates, with only a relatively small (roughly 10-20 MK) increase in temperature required to fall in the acceptable band of parameter space identified. The other potential candidate identified was Novae, although the lack of detailed models of white dwarfs accreting material similar to NGC 2419 either from a binary companion or the intra cluster medium limit appraisals of its feasibility regarding whether enough ejecta could be produced and retained by the cluster. However, based on current novae frequency in globular clusters determined by \cite{kato2013novae}, \cite{iliadis2016} conclude that the amount of material that would be produced by Novae is at most 1 \% of the total required mass to pollute 30 \% of NGC 2419.

@@

The data set of 454180 LAMOST giants was prepared by \citet{ho2017} for use with the supervised machine learning code \textit{The Cannon}. The LAMOST data was shifted to rest frame, normalised, and sampled to a common uniform wavelength grid between 3905 \AA \\ and 9000 \AA. \textit{The Cannon} was used to infer $\text{T}_{\text{eff}}$, log(g), [Fe/H] and [$\alpha$/M] using 9952 common stars between LAMOST and the higher resolution APOGEE survey through the use of a data-driven spectral model, referred to henceforth and simply 'the model'. For details regarding the preparatory work, model generation and label transfer between the APOGEE and LAMOST surveys, the reader is directed to \citet{ho2017}. The data-driven model generated by \textit{The Cannon} is treated in this study as the typical spectra for each star.

@@

Three filters were applied independently, with the candidate pool formed from all unique stars identified each of the three filters. The first filter required the Mg 5184 line's Gaussian to satisfy $Amp > 0.05$ and $\frac{Amp}{Amp_{err}}>3$, and the K 7665 line to satisfy $Amp < 0.05$ and $\frac{Amp}{Amp_{err}}>3$. The second, more demanding filter was applied to require 2 of the three Mg lines to satisfy $Amp > 0.05$ and $\frac{Amp}{Amp_{err}}>3$, as well as either of the two K lines to satisfy $Amp < 0.05$ and $\frac{Amp}{Amp_{err}}>3$. The third filter removed the amplitude requirement to requiring only positivity on 2/3 Mg lines and negativity on \textit{both} K lines, in addition to data quality requirements $\chi^2<3$ and $S/N>30$.
These filters together produced 384 unique candidate stars to be vetted manually.

The vetting process was done through visual examination of each case. The minimum criterion for a pass was two out of the three Mg lines to look at least plausibly depleted, and at least one of the K lines to look convincing. The apparent noisiness of the signal was qualitatively taken into account when making these determinations. Of the 384 candidates manually vetted, 188 were ultimately determined to be sufficiently strong to be reported. The 6556 \AA \\ H line was then visually examined for each star to identify stars with active chromosphere emissions which could result in false positives for Mg depletion. 62 stars were removed from the sample due to being chromospheric emmiters. A further 13 stars were removed due to the model fitted poorly to the H 6556 line possibly due to incorrect log(g) or $\text{T}_{\text{eff}}$ estimates from \textit{The Cannon}, leaving a total of 113 candidates displaying depleted Mg and enhanced K.

@@

\begin{figure}
	\includegraphics[width=\columnwidth]{posterchildof13.png}
    \caption{Poster Child}
    \label{mhist}
\end{figure}

\begin{figure}
	\includegraphics[width=\columnwidth]{histof113.png}
    \caption{Metalicity Histogram}
    \label{mhist}
\end{figure}

\begin{figure}
	\includegraphics[width=\columnwidth]{loggteffof113.png}
    \caption{Effective temperature, log surface gravity, metalicity}
    \label{mhist}
\end{figure}

\begin{figure}
	\includegraphics[width=\columnwidth]{globclustof113.png}
    \caption{Galactic coordinates for Mg-K stars and known globular clusters}
    \label{mhist}
\end{figure}
